---
title             : "The influence of confidence on Judgments of Associate Memory"
shorttitle        : "Confident JAM"

author: 
  - name          : "Erin M. Buchanan"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "901 S. National Ave, Springfield, MO, 65897"
    email         : "erinbuchanan@missouristate.edu"
  - name          : "Amber Gillenwaters"
    affiliation   : "1"

affiliation:
  - id            : "1"
    institution   : "Missouri State University"
  
author_note: |
  Erin M. Buchanan is an Associate Professor of Psychology at Missouri State University, and Amber Gillenwaters is a masters degree candidate in Experimental Psychology at Missouri State University.
  
abstract: |
  Enter abstract here. Each new line herein must be indented, like this line.
  
keywords          : "keywords"

#bibliography      : ["r-references.bib"]

figsintext        : no
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes
mask              : no

class             : "man"
output            : papaja::apa6_pdf
---

```{r load_packages, include = FALSE}
library("papaja")

#load the data
master = read.csv("complete_data.csv")
```

don't run this block, I already made the data
```{r create-data, include = FALSE, eval = FALSE}
##this code shows how we combined the data
#set working directory to the data folder
temp = list.files(pattern="*.dat")
myfiles = lapply(temp, read.delim)
dataset = do.call(rbind, myfiles)

dataset$partno = with(dataset, paste(date,time,subject, sep = ""))
table(dataset$partno)

wordpairs = read.delim("finalpairs2.txt", header = F)
wordpairs$index = paste(tolower(as.character(wordpairs$V2)),
                        tolower(as.character(wordpairs$V3)), sep = ".") 

##these files can be found in our SPP or wordnorms2 github
swow = read.csv("strength.SWOW-EN.R123.csv")
swow$index = paste(tolower(as.character(swow$cue)),
                   tolower(as.character(swow$response)), sep = ".")
usf = read.delim("usf_norms.txt")
usf$index = paste(tolower(as.character(usf$CUE)),
                  tolower(as.character(usf$TARGET)), sep = ".")

library(expss)

wordpairs$fsg = vlookup(wordpairs$index, usf, "FSG", "index")
wordpairs$bsg = vlookup(wordpairs$index, usf, "BSG", "index")
wordpairs$swow = vlookup(wordpairs$index, swow, "R123.Strength", "index")

dataset$fsg = vlookup(dataset$trialcode, wordpairs, "fsg", "V1")
dataset$bsg = vlookup(dataset$trialcode, wordpairs, "bsg", "V1")
dataset$swow = vlookup(dataset$trialcode, wordpairs, "swow", "V1")  
write.csv(dataset, "complete_data.csv", row.names = F)
```


# Method

## Participants

## Material

## Procedure

## Data analysis

# Results

```{r data-screening, include = FALSE}
##remove text responses from the response column
master$response = as.numeric(as.character(master$response))

##remove those lines from the data

##remove practice trials

##only screen the response column

#rescale responses to 0 to 100
master$response = master$response * 10 - 5

#rescale others so all 100 scale
master[ , c("fsg", "bsg", "swow")] = master[ , c("fsg", "bsg", "swow")] * 100
```

Hypothesis 1: 
The JAM effect will replicate previous research.

Analyze with MLM:
Use partno as random factor
Test if trialcode needs to be second random factor
Use FSG as IV
Use response as the DV - subset the data without the confidence trials

```{r mlm-hyp1, include = FALSE}

```

Hypothesis 2:
Does confidence change within the course of the experiment?

Analyze with MLM:
Use partno as random factor
Use trialnum as the IV
Use response as the DV - subset the data to only confidence trials

```{r mlm-hyp2, include = FALSE}

```

Hypothesis 3:
Does confidence relate to the actual JAM? 
- Figure out the difference score between FSG and response (level of accuracy) for the trial before confidence 
- Use confidence to predict that score

Analyze with MLM:
Use partno as random factor
Use confidence*fsg as the IV
Use response as the DV 

```{r mlm-hyp3, include = FALSE}
##creating a dataset that will work for you
confidencetrials = seq(5,120,5)
jamtrials = seq(4,119,5)

confidencedata = master[master$trialnum %in% confidencetrials , ]
judgmentdata = master[master$trialnum %in% jamtrials, ]

##this assumes they are lined up the same, I will need to check it after data screening
judgmentdata$confidence = confidencedata$response

```


# Discussion


\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup
